---
title: "p8105_hw2_sl5009"
author: "Shengzhi Luo"
date: "08/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggridges)
```

## Prolem 1
Read and clean the Mr.Trash Wheel sheet.
```{r,message=FALSE}
Mr_Trash_Wheel<-read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx")
Mr_Trash_Wheel=janitor::clean_names(Mr_Trash_Wheel)
```

Omit rows that do not include dumpster-specific data.
```{r,results=FALSE}
Mr_Trash_Wheel=drop_na(Mr_Trash_Wheel,dumpster)
```

Round the number of sports balls to the nearest integer.
```{r,results=FALSE}
Mr_Trash_Wheel$sports_balls=round(Mr_Trash_Wheel$sports_balls,digits = 0)
```

Read and clean precipitation data for 2018 and 2019.
```{r,message=FALSE}
Pre_2018=
  read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet="2018 Precipitation")%>%
  janitor::clean_names()%>%
  drop_na()%>%
  mutate(
    year=2018
    )
Pre_2018=Pre_2018[-1,]
```


```{r,message=FALSE}
Pre_2019=
  read_excel("./Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet="2019 Precipitation")%>%
  janitor::clean_names()%>%
  drop_na()%>%
  mutate(year=2019)
Pre_2019=Pre_2019[-1,]
```

Combine precipitation datasets and convert month to a character variable. 
```{r}
Pre_2018_2019=
  bind_rows(Pre_2018,Pre_2019) %>%
  mutate(
    precipitation_in=month.name[as.numeric(precipitation_in)]
  )
Pre_2018_2019
```

```{r}
Pre_2018_sum=sum(as.numeric(Pre_2018$x2))
```

```{r}
Mr_Trash_Wheel_2019=filter(Mr_Trash_Wheel,year=="2019")
Mr_Trash_Wheel_median=median(Mr_Trash_Wheel_2019$sports_balls)
```

The precipitation data on May, July and September of 2018 are very rainy than the rest of the time of 2018 while data on 2019 are relatively low comparing to 2018 throughout the whole year of 2019. The total precipitation in 2018 is `r Pre_2018_sum` and the median number of sports balls in a dumpster in 2019 is `r Mr_Trash_Wheel_median`.

## Problem 2

Clean the data in pols-month.csv.
```{r}
pols_month=
  read.csv("./fivethirtyeight_datasets/pols-month.csv")%>%
  separate(mon,into=c("year","month","day"))%>%
  mutate(
    president=prez_gop+prez_dem,
    month = month.abb[as.numeric(month)],
    year=as.numeric(year))%>%
  select(-c("prez_gop","prez_dem","day"))
```


Clean the data in snp.csv.
```{r}
snp=
  read.csv("./fivethirtyeight_datasets/snp.csv")%>%
  separate(date,into=c("month","day","year"))%>%
  arrange(as.numeric(month))%>%
  mutate(
    month = month.abb[as.numeric(month)]
    )%>%
  select(-c("day"))%>%
  relocate(year,month,close)%>%
  mutate(
    year=as.numeric(year),
    year=ifelse(year>47, 1900+year, 2000+year)
  )%>%
  arrange(year)
```

Tidy the unemployment data so that it can be merged with the previous datasets.
```{r}
unemployment=
  read.csv("./fivethirtyeight_datasets/unemployment.csv")%>%
  pivot_longer(
    Jan:Dec,
    names_to="month",
    values_to = "unemployment rate"
  )
colnames(unemployment) <- c("year","month","unemployment rate")
```

Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r}
fas_data=
  left_join(pols_month,snp,by=c("year","month"))
res_data=
  left_join(fas_data,unemployment,by=c("year","month"))
```

The pols_month dataset contained 822 observations of 9 variables related to the number of national politicians who are democratic or republican at any given time as governors, senators, representatives and president.
The snp dataset contained tandard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole with the closing values of the S&P stock index on each months from 1950-2015.
The unemployment dataset contained percentage of unemployment in each months from 1948-2015.

## Problem 3
Load and tidy the data.
```{r}
Popular_Baby_Names=
  read.csv("./Popular_Baby_Names.csv")%>%
  janitor::clean_names() %>%
  mutate(child_s_first_name=stringr::str_to_title(child_s_first_name))%>%
  mutate(ethnicity=replace(ethnicity,ethnicity=="ASIAN AND PACI","ASIAN AND PACIFIC ISLANDER"),
         ethnicity=replace(ethnicity,ethnicity=="BLACK NON HISP","BLACK NON HISPANIC"),
         ethnicity=replace(ethnicity,ethnicity=="WHITE NON HISP","WHITE NON HISPANIC")
         )
Popular_Baby_Names=distinct(Popular_Baby_Names)
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time.
```{r}
Olivia_rank=
  filter(Popular_Baby_Names, child_s_first_name=="Olivia")%>%
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = c(count,rank)
  )%>%
  relocate(child_s_first_name,ethnicity)
```

Produce a table showing the most popular name among male children over time.
```{r}
male_rank=
  filter(Popular_Baby_Names, gender=="MALE")%>%
  filter(rank==1)%>%
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = c(count,rank)
  )
```

Produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).
```{r}
male_white_2016=
  filter(Popular_Baby_Names,gender=="MALE",ethnicity=="WHITE NON HISPANIC",year_of_birth=="2016")
male_white_2016%>%
  ggplot(aes(x=rank,y=count))+
  geom_point()
```




